### 2018-04-25 manuscript.rmd
updated tree stats

Major questions remained after this early work, though. In particular, there was little support for important deep relationships within Codonophora. Understanding these relationships is key to resolving the evolution of several traits of importance, including sexual systems (monoecy versus dioecy) and the gain and loss of particular zooids, such as palpons (Fig. 2). Here we present a broadly sampled phylogenetic analysis of Siphonophora that considers transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were subsequently excluded due to poor sampling). Using 1,071 genes, we find strong support for many relationships found in the earlier phylogeny [@Dunn:2005dy], and also provide new resolution for key relationships that were unresolved in that previous study. Using this phylogeny, we reconstruct the evolutionary history of characters central to the unique biology of siphonophores, including zooid type, life history traits, and habitat.

@@

New data were analysed in conjunction with 14 publicly available datasets [@zapata2015phylogenomic; @Dunn:2013kw; @lehnert2012developing; @philippe2009phylogenomics; @fidler2014unique; @chapman2010dynamic; @sanders2014differential; @putnam2007sea; @sanders2015interspecific], with a total number of 43 species. Sequence assembly, annotation, homology evaluation, gene tree construction, parsing of genes trees to isolate orthologous sequences, and supermatrix construction were conducted with Agalma v. 1.0.0 [@Dunn:2013kw; @Guang202416].  This workflow integrates a variety of existing tools [@grabherr2011full; @altschul1990basic; @doi:10.1093/nar/30.7.1575; @sukumaran2010dendropy; @langmead2012fast; @li2011rsem; @talavera2007improvement; @li2009sequence; @katoh2013mafft] and new methods. Maximum likelihood (ML) analyses of the supermatrix were conducted with RAxML v 8.2.0 [@stamatakis2006raxml] and implemented via Agalma. We also used IQTree to implement other ML analyses using different models of substitution to test the robustness of the results to model selection [@doi:10.1093/molbev/msu300; @doi:10.1093/molbev/msx281]. Bayesian Inference (BI) analyses of the supermatrix were conducted using Phylobayes v. 1.7a-mpi [@lartillot2009phylobayes]. Sequence alignments, sampled and consensus trees, and voucher information are available in the git repository. Tree figures were rendered with ggtree [@Yu:2016fo].

Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix due to low gene occupancy (gene sampling of 20.8% and 14.5% respectively in a 50% occupancy matrix with 2,203 genes). The final analyses presented here consider 33 siphonophore species and 8 outgroup species. This includes new data for 30 species. In the final analyses, we sampled 1,071 genes to generate a supermatrix with 60% occupancy and a length of 378,468 amino acids (Fig. S1).

ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\Gamma$ model of amino acid substitution, and bootstrap values were estimated using `r length(bootstrap_trees)` replicates. BI was conducted using two different CAT models, CAT-Poisson and CAT-GTR [@doi:10.1093/molbev/msh112]. Two independent MCMC chains were run under the CAT-GTR model, and four independent MCMC chains were run under the CAT-Poisson model. The BI analyses did not converge (CAT-Poisson: maxdiff=1, meandiff=0.0148; CAT-GTR: maxdiff=1, meandiff=0.0127). Only the results from the CAT-Poisson model are presented here. Visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all CAT-Poisson runs. This left `r length( phylobayes_trees )` trees in the posterior.

@@

The phylogenetic relationships recovered in this study received strong support across analysis methods (Fig. 3A), with a couple of localized exceptions (Fig. 3B). The ML results were robust to model selection (Fig. S10). Most clades are consistent with those found in a previous study based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy]. Relationships that receive strong support in both include the placement of Cystonectae as sister to Codonophora (the clade that includes all other siphonophores), the placement of Apolemiidae as sister to all other codonophorans, and the placement of Calycophorae within the paraphyletic "Physonectae". Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in the present 1,071-gene transcriptome analyses. There is strong support for Pyrostephidae as sister to all other non-apolemiid codonophorans. We provisionally refer here to Pyrostephidae as the clade including *Rudjakovia* sp., although sampling of *Pyrostephos vanhoeffeni* is needed in order to determine if *Rudjakovia* sp. falls within Pyrostephidae or is sister to it. Within the clade that is sister to Pyrostephidae, we find two main clades, Calycophorae and a clade we here name Euphysonectae (Fig. 3A). It includes the remaining non-apolemiid, non-pyrostephid "Physonectae". We define Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to *Diphyes dispar*.

@@

![60% gene occupancy matrix for 41 species across 1,071 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/Supplementary_Figure1.png)

@@

```{r supplementary trees, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Phylogenies obtained using different models of evolution. Unlabeled nodes have support >0.99. A. GTR + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories. Likelihood: -5964623.1967; AIC score: 11929462.3933.  B. JTT + Empirically counted frequencies from alignment + FreeRate model with 6 categories. Likelihood:-5964771.2358; AIC: 11929758.4715. C. WAG + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories. Likelihood:-5985124.7530; AIC score: 11970465.5060 "}

>>>>

Major questions remained after this early work, though. In particular, there was little support for important deep relationships within Codonophora. Understanding these relationships is key to resolving the evolution of several traits of importance, including sexual systems (monoecy versus dioecy) and the gain and loss of particular zooids, such as palpons (Fig. 2). Here we present a broadly sampled phylogenetic analysis of Siphonophora that considers transcriptomic data from 33 siphonophore species and 10 outgroup species (2 outgroups were subsequently excluded due to poor sampling). Using 1,423 genes, we find strong support for many relationships found in the earlier phylogeny [@Dunn:2005dy], and also provide new resolution for key relationships that were unresolved in that previous study. Using this phylogeny, we reconstruct the evolutionary history of characters central to the unique biology of siphonophores, including zooid type, life history traits, and habitat.

@@

New data were analysed in conjunction with 14 publicly available datasets [@zapata2015phylogenomic; @Dunn:2013kw; @lehnert2012developing; @philippe2009phylogenomics; @fidler2014unique; @chapman2010dynamic; @sanders2014differential; @putnam2007sea; @sanders2015interspecific], with a total number of 43 species. Sequence assembly, annotation, homology evaluation, gene tree construction, parsing of genes trees to isolate orthologous sequences, and supermatrix construction were conducted with Agalma (commit 6bd9988, running BioLite commit 784edc6) [@Dunn:2013kw; @Guang202416].  This workflow integrates a variety of existing tools [@grabherr2011full; @altschul1990basic; @doi:10.1093/nar/30.7.1575; @sukumaran2010dendropy; @langmead2012fast; @li2011rsem; @talavera2007improvement; @li2009sequence; @katoh2013mafft] and new methods. Maximum likelihood (ML) analyses of the supermatrix were conducted with RAxML v 8.2.0 [@stamatakis2006raxml] and implemented via Agalma. We also used IQTree to implement other ML analyses using different models of substitution to test the robustness of the results to model selection [@doi:10.1093/molbev/msu300; @doi:10.1093/molbev/msx281]. Bayesian Inference (BI) analyses of the supermatrix were conducted using Phylobayes v. 1.7a-mpi [@lartillot2009phylobayes]. Sequence alignments, sampled and consensus trees, and voucher information are available in the git repository. Tree figures were rendered with ggtree [@Yu:2016fo].

Two outgroup species, *Atolla vanhoeffeni* and *Aegina citrea*, were removed from the final supermatrix due to low gene occupancy (gene sampling of 17.0% and 17.3% respectively in a 60% occupancy matrix with 3,379 genes). The final analyses presented here consider 33 siphonophore species and 8 outgroup species. This includes new data for 30 species. In the final analyses, we sampled 1,423 genes to generate a supermatrix with 80% occupancy and a length of 395,699 amino acids (Fig. S1).

ML analyses were conducted on the unpartitioned supermatrix using the WAG+$\Gamma$ model of amino acid substitution, and bootstrap values were estimated using `r length(bootstrap_trees)` replicates. BI was conducted using two different CAT models, CAT-Poisson and CAT-GTR [@doi:10.1093/molbev/msh112]. Two independent MCMC chains were run under the CAT-GTR model, and eight independent MCMC chains were run under the CAT-Poisson model. The BI analyses did not converge (CAT-Poisson: maxdiff=1, meandiff=0.0148; CAT-GTR: maxdiff=1, meandiff=0.0221709). Only the results from the CAT-Poisson model are presented here. Two chains (chain1 and chain3) did not burn in after 1405 and 4695 iterations and were excluded from the analyses. Visual inspection of the traces indicated that a burn in of `r burn_in` trees was sufficient for all CAT-Poisson runs. This left `r length( phylobayes_trees )` trees in the posterior.

@@

The phylogenetic relationships recovered in this study received strong support across analysis methods (Fig. 3A), with a couple of localized exceptions (Fig. 3B). The ML results were robust to model selection (Fig. S10). Most clades are consistent with those found in a previous study based on two genes (16S and 18S ribosomal RNA) [@Dunn:2005dy]. Relationships that receive strong support in both include the placement of Cystonectae as sister to Codonophora (the clade that includes all other siphonophores), the placement of Apolemiidae as sister to all other codonophorans, and the placement of Calycophorae within the paraphyletic "Physonectae". Multiple nodes that were not resolved in the previous two-gene analysis receive strong support in the present 1,423-gene transcriptome analyses. There is strong support for Pyrostephidae as sister to all other non-apolemiid codonophorans. We provisionally refer here to Pyrostephidae as the clade including *Rudjakovia* sp., although sampling of *Pyrostephos vanhoeffeni* is needed in order to determine if *Rudjakovia* sp. falls within Pyrostephidae or is sister to it. Within the clade that is sister to Pyrostephidae, we find two main clades, Calycophorae and a clade we here name Euphysonectae (Fig. 3A). It includes the remaining non-apolemiid, non-pyrostephid "Physonectae". We define Euphysonectae as the clade consisting of *Agalma elegans* and all taxa that are more closely related to it than to *Diphyes dispar*.

@@

![60% gene occupancy matrix for 41 species across 1,423 genes. Genes and species are sorted by sampling, the best sampled shown in the upper left.](supplementary_info/Supplementary_Figure1.png)

@@

```{r supplementary trees, echo=FALSE, message=FALSE, warning=FALSE, verbose= FALSE, fig.width=6.5, fig.height=7, fig.cap= "Phylogenies obtained using different models of evolution. Unlabeled nodes have support >0.99. A. GTR + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories. Likelihood: -8133157.335; AIC:16266530.7277.  B. JTT + Empirically counted frequencies from alignment + FreeRate model with 7 categories. Likelihood:-8113694.922; AIC: . C. WAG + Optimized base frequencies by maximum-likelihood + Free rate model with 6 categories. Likelihood: -8156043.772; AIC score: 16312303.6120 "}
